---
title: '<span style=''color: #228B22;''>University of San Francisco</span><br><span style=''color: #CDAD00;''>Women''s Basketball
  Team Scouting Report</span>'
author: ''
output:
  html_document: default
---

<style>
h1.title {
  text-align: center;
}
</style>

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE}
setwd("~/Downloads/WBB Scouting")

# Comparison of Teams, Game by Game
games = read.csv("LMU Data - Games.csv")
library(mosaic)
teams <- levels(factor(games$Team))

# Comparison of Teams, Distribution of Workload Among Players

wcc_players = read.csv("WCC Players - WCCData.csv")
```

```{r, echo = FALSE}
# Four Factors

# 1. Effective Field Goal Percentage
games$eFG = (games$FG + (0.5 * games$X3P)) / games$FGA
games$DeFG = (games$DFG + (0.5 * games$D3P)) / games$DFGA

# 2. Free Throw Rate
games$Ftr = games$FTA / games$FG
games$DFtr = games$DFTA / games$DFG

# 3. Rebounding Percentage

games$RBpct = games$ORB / (games$ORB + games$DDRB)
games$DRBpct = games$DRB / (games$DORB + games$DRB)

# 4. Turnover Rate

games$TOVpct = games$TOV / (games$FGA + (0.44 * games$FTA) + games$TOV)
games$DTOVpct = games$DTOV / (games$DFGA + (0.44 * games$DFTA) + games$DTOV)

```

$\\$

# Overall Key Points

$\\$

- Medicore results on offense

- Too many fouls on defense

- Even split of defensive rebounding

$\\$

# Offense

$\\$

**<span style="color: green;">Strengths:</span>** none in particular



**<span style="color: red;">Weaknesses:</span>** none in particular

```{r, echo = FALSE}
# eFG

means = mean(eFG ~ Team, data = games)

ses = tapply(games$eFG, games$Team,
             function(x) sd(x)/sqrt(length(x)))


dotchart(means, xlim = c(0.38, 0.54),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Offensive eFG",
         xlab = "eFG%")
segments(means - ses, 1:12, 
         means + ses, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))

# FTr

means_ftr = mean(Ftr ~ Team, data = games)

ses_ftr = tapply(games$Ftr, games$Team,
                  function(x) sd(x)/sqrt(length(x)))

dotchart(means_ftr, xlim = c(0.45, 0.88),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Free Throw Ratio",
         xlab = "FT / FGA")
segments(means_ftr - ses_ftr, 1:12, 
         means_ftr + ses_ftr, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))

# ORB Pct

means_rbpct = mean(RBpct ~ Team, data = games)

ses_rbpct = tapply(games$RBpct, games$Team,
                  function(x) sd(x)/sqrt(length(x)))

dotchart(means_rbpct, xlim = c(0.19, 0.43),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Offensive Rebound Percentage in WCC Play",
         xlab = "ORB%")
segments(means_rbpct - ses_rbpct, 1:12, 
         means_rbpct + ses_rbpct, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))

# TOV Pct
means_tovpct = mean(TOVpct ~ Team, data = games)

ses_tovpct = tapply(games$TOVpct, games$Team,
                    function(x) sd(x)/sqrt(length(x)))

dotchart(means_tovpct, xlim = c(0.1, 0.25),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Turnover Percentage",
         xlab = "TOV%")
segments(means_tovpct - ses_tovpct, 1:12, 
         means_tovpct + ses_tovpct, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))
```




# Defense

$\\$

**<span style="color: green;">Strengths:</span>** none in particular

**<span style="color: red;">Weaknesses:</span>** too few forced
turnovers, too many fouls

```{r, echo = FALSE}
# DeFG

means_defg = mean(DeFG ~ Team, data = games)

ses_defg = tapply(games$DeFG, games$Team,
             function(x) sd(x)/sqrt(length(x)))

dotchart(means_defg, xlim = c(0.38, 0.54),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Opponent eFG",
         xlab = "eFG% Allowed")
segments(means_defg - ses_defg, 1:12, 
         means_defg + ses_defg, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))


# DFtr

means_dftr = mean(DFtr ~ Team, data = games)

ses_dftr = tapply(games$DFtr, games$Team,
                 function(x) sd(x)/sqrt(length(x)))

dotchart(means_dftr, xlim = c(0.35, 0.95),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Opponent Free Throw Ratio",
         xlab = "FT / FGA Allowed")
segments(means_dftr - ses_dftr, 1:12, 
         means_dftr + ses_dftr, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))

# DRB Pct

means_drbpct = mean(DRBpct ~ Team, data = games)

ses_drbpct = tapply(games$RBpct, games$Team,
                   function(x) sd(x)/sqrt(length(x)))

dotchart(means_drbpct, xlim = c(0.63, 0.8),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Defensive Rebound Percentage Comparison",
         xlab = "DRB%")
segments(means_drbpct - ses_drbpct, 1:12, 
         means_drbpct + ses_drbpct, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))

# DTOV pct

means_dtovpct = mean(DTOVpct ~ Team, data = games)

ses_dtovpct = tapply(games$DTOVpct, games$Team,
                    function(x) sd(x)/sqrt(length(x)))

dotchart(means_dtovpct, xlim = c(0.12, 0.25),
         col = ifelse(teams == "San Francisco", "forestgreen", "black"),
         main = "Opponent Turnover Percentage",
         xlab = "TOV%")
segments(means_dtovpct - ses_dtovpct, 1:12, 
         means_dtovpct + ses_dtovpct, 1:12,
         col = ifelse(teams == "San Francisco", "gold3", "black"))
```



# Workload Distribution

```{r, echo = FALSE}
# Shannon Diversity Index

Shannon_div = function(vec){
  sum = 0
  for(i in 1:length(vec)){
    if(vec[i] == 0){
      sum = sum + 0
    } else{
    sum = sum + (vec[i] * log(vec[i]))
    }
  }
  return(-1 * sum)
}

# Simpson's Diversity Index

Simpsons_div = function(vec){
  N = sum(vec)
  sum = 0
  for(i in 1:length(vec)){
    sum = sum + (vec[i] * (vec[i] - 1))
  }
  
  denom = N * (N - 1)
  
  return(1 - (sum / denom))
}
```

```{r, echo = FALSE}
# Data Analysis

# Initializing Data

divs_pts_Shan = c()
divs_pts_Simpsons = c()

divs_ast_Shan = c()
divs_ast_Simpsons = c()

divs_drb_Shan = c()
divs_drb_Simpsons = c()

# Looping Through 12 Teams to Calculate Diversity Indices

for(i in 1:12){
  school = unique(wcc_players$Team)[i]
  
  # Points
  
  school_pts_tot = wcc_players$PTS[wcc_players$Team == school]
  
  divs_pts_Shan = c(divs_pts_Shan,
                    Shannon_div(school_pts_tot / sum(school_pts_tot)))
  
  divs_pts_Simpsons = c(divs_pts_Simpsons,
                        Simpsons_div(school_pts_tot))
  
  # Assists
  
  school_ast_tot = wcc_players$AST[wcc_players$Team == school]
  
  divs_ast_Shan = c(divs_ast_Shan,
                    Shannon_div(school_ast_tot / sum(school_ast_tot)))
  
  divs_ast_Simpsons = c(divs_ast_Simpsons,
                        Simpsons_div(school_ast_tot))
  
  # DRB
  
  school_drb_tot = wcc_players$DRB[wcc_players$Team == school]
  
  divs_drb_Shan = c(divs_drb_Shan,
                    Shannon_div(school_drb_tot / sum(school_drb_tot)))
  
  divs_drb_Simpsons = c(divs_drb_Simpsons,
                        Simpsons_div(school_drb_tot))
}

# Labeling Vectors

names(divs_pts_Shan) <- unique(wcc_players$Team)
names(divs_pts_Simpsons) <- unique(wcc_players$Team)

names(divs_ast_Shan) <- unique(wcc_players$Team)
names(divs_ast_Simpsons) <- unique(wcc_players$Team)

names(divs_drb_Shan) <- unique(wcc_players$Team)
names(divs_drb_Simpsons) <- unique(wcc_players$Team)
```

### Scoring

$\\$

**Assessment:** inconclusive

```{r, echo = FALSE}
# Points

x = sort(divs_pts_Shan)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Shannon Diversity Index",
     main = "Distribution of Scoring - Shannon",
     ylim = c(2, 2.45),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))
     

# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx],x[sc_idx] + 0.05,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")


x = sort(divs_pts_Simpsons)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Simpson's Diversity Index",
     main = "Distribution of Scoring - Simpson's",
     ylim = c(0.84, 0.91),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))


# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx] - 0.3, x[sc_idx] + 0.005,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")
```

### Assisting

$\\$

**Assessment:** inconclusive

```{r, echo = FALSE}
# Assists

x = sort(divs_ast_Shan)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Shannon Diversity Index",
     main = "Distribution of Assists - Shannon",
     ylim = c(1.7, 2.3),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))


# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx] - 0.2,x[sc_idx] + 0.06,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")

x = sort(divs_ast_Simpsons)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Simpson's Diversity Index",
     main = "Distribution of Assists - Simpson's",
     ylim = c(0.75, 0.88),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))


# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx] - 0.2,x[sc_idx] + 0.02,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")
```

### Defensive Rebounding

$\\$

**Assessment:** pretty strong distribution of defensive rebounding

```{r, echo = FALSE}
# DRB

x = sort(divs_drb_Shan)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Shannon Diversity Index",
     main = "Distribution of Defensive Rebounding - Shannon",
     ylim = c(1.9, 2.5),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))


# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx],x[sc_idx] + 0.085,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")

x = sort(divs_drb_Simpsons)
y = 1:12

plot(y, x,
     pch = 19,
     xaxt = "n",
     ylab = "Simpson's Diversity Index",
     main = "Distribution of Defensive Rebounding - Simpson's",
     ylim = c(0.82, 0.91),
     xlab = "WCC Team",
     xlim = c(1, length(x)),
     col = ifelse(names(x) == "San Francisco", "gold3", "black"))

segments(x0 = y, y0 = 0,
         x1 = y, y1 = x,
         col = ifelse(names(x) == "San Francisco", "gold3", "black"))


# label only San Francisco
sc_idx <- which(names(x) == "San Francisco")

text(y[sc_idx] - 0.2,x[sc_idx] + 0.015,
     labels = "San Francisco",
     cex = 1.15,
     col = "forestgreen")
```



# Appendix

$\\$

**Effective Field Goal Percentage:**

$$eFG = \frac{FG + \frac{1}{2}3P}{FGA}$$

$3P$: three point shots made

$\\$

**Free Throw Percentage:**

$$FTR = \frac{FTA}{FGA}$$


$\\$

**Rebounding Percentage:**

$$RB_{pct}  = \frac{ORB}{ORB+Opp \ DRB}$$



flip $ORB$, $DRB$ for defensive rebounding metric

$\\$

**Turnover Percentage:**

$$TOV_{pct} = \frac{TOV}{FGA + TOV + \frac{11}{25}FTA}$$

$\\$

**Shannon's Diversity Index:**

$H = - \underset{i \ \in \ C}{\sum} \ (p_i\log(p_i))$,

$p_i$ is proportion of $C$ that belongs to $i$

$\\$

**Simpson's Diversity Index:**

$D = 1 - \frac{\underset{i \ \in \ C}{\sum} \ N_i(N_i-1)}{N(N-1)}$,

$N$ is the total number of elements in $C$, $N_i$ is
the number of elements in $C$ that belong to group $i$
